### 2021-12-16 documentos/generators.tex
Otra iteración de sección Generators lista

In this section we will discuss the existence of a generator set for all PCE
quantum channels, how to label each of them uniquely according to its local action 
on every qubit in the system, and, finally, we  	will discuss a symmetry of
both PCE quantum channel generators and $A$, the matrix that diagonalizes
the Choi matrix $\mcD_N$ of a PCE map.

There exists a subset of PCE channels that generate 
all PCEs; the nature of these generators may be studied, as we
shall see, with the properties of the aforementioned vector space.
By standard theorems of linear algebra, any proper subspace $W$, see \Eref{eq:19}, 
can be extended to a maximal non-trivial subspace of dimension $2N-1$ by
adjoining appropriate additional basis elements. This can be done in
different ways. We therefore arrive to the set of maximal extensions of $W$,
where every maximal subspace corresponds to a PCE quantum channel that
preserves half of the Pauli components. The intersection of all the elements of
this set reduces to $W$ itself,
and since intersection of subspaces translates to composition of PCEs,
this implies that all PCEs can be obtained as
compositions of PCEs corresponding to maximally non-trivial subspaces, plus
the identity map. In other words, the set of PCEs that preserve half 
of the components plus the identity map, are a generator set for 
all PCE channels.
% and the complete vector space (corresponding to the identity channel),
% which we call generators. 
% \cpnote{Aca tenemos un semigrupo y no se si incluir a la identidad como 
% generador. De alguna manera es excepcional. Es problematico incluirla y es 
% problematico no incluirla.}
Consider \Fref{fig:examplesPCE}; subfigures (c), (d) and (e) represent nontrivial 
PCE generators and the composition of any two of them yield the PCE channel
corresponding to (b). 
% 
% 
% the composition of any two 2-qubits PCE
% generators 
% in (c), (d), or (e) is equivalent to taking the intersection between them both
% and the three different possible compositions result in PCE channel in (b).
%\cpnote{Aca se me hace que hay un cambio brusco de tema. Ver que onda
%con Jose Alfredo}
%The existence of the PCE generators $\mcG$ provide the program of 
%constructing all PCE channels a tool that avoids the usage of a 
%high-computational-cost method. In a moment, we will discuss how to find the 
%generators $\mcG$ from matrix $A$, given a symmetry that is obeyed both by the generators 
%$\mcG$ and matrix $A$ in \Eref{eq:san}\cpnote{Como que para que ahora se
%menciona entonces?}.
%For the case of 1 qubit, the existence of PCE generators $\pceg$ is  
%verified by Siudzińska~\cite{siudzinska2019geometry}.
%\janote{Esto puede
%ser un párrafo, o puede ser un comentario final dentro del párrafo antrerior.}

%\cpnote{Arreglar}
PCE generators may be characterized by its local action on every qubit 
in the system, and thus each of them be denoted as $\mcG_{\vec\alpha}$, 
with $\vec\alpha$ each of the $4^{n}$ different vector indices. For 1-qubit
the PCE generators are the identity map plus the channels that map the Bloch 
sphere to each Cartesian axis, as their associated vector subspaces are trivial
to prove and their dimension 1 is met. 
\janote{Here goes the bla bla for the discussion of the proof for the $N$-qubit
\pce{} generators}
Therefore, we can identify the index $\alpha_k$ in $\vec\alpha$ by the local action
of the PCE generator on the $k$-th qubit. 
See, for example, that the 2-qubits PCE generator represented in subfigure (c)
of \Fref{fig:examplesPCE} acts on the first qubit (first column) as mapping its
Bloch sphere to de $x$-axis, and on the second qubit (first row) acts an identity,
hence it is labeled $\mcG_{(1,0)}$.
See \Fref{fig:appex:twoQ_PCEGs} for the notation of all 2-qubit PCE generators.

%\cpnote{Mucha lora y no se prueba nada}
A reflection symmetry is identified for both PCE generators $\pceg$ and
the matrix $A$ that diagonalizes the Choi matrix $\mcD_N$ of a PCE map.
%PCE generators $\pceg$ obey a symmetry
%that defines uniquely every PCE generator. 
The symmetry for a PCE generator $\pceg$ is encoded in the 
parameters $\{\taus\}$, the $N$-dimensional Cartesian grid associated with it 
(the set) is either symmetric or anti-symmetric under reflection of the 
indices in the $k$-th position in $\vec\alpha$ as
\begin{align}
\alpha_{0}\mapsto\alpha_3,\quad\alpha_{3}\mapsto\alpha_0,
\quad\alpha_1\mapsto\alpha_2, \quad\alpha_2\mapsto\alpha_1.
\end{align}
the $N$ hyperplanes of dimension $N-1$ that divide the grid in half.
For instance, see that the diagrams representing the generators 
$\mcG_{(1,3)}$, $\mcG_{(3,2)}$ and $\mcG_{(2,2)}$ in subfigures 
(c), (d) and (e) of \Fref{fig:two:qubit:examples}, respectively,
are either symmetric or anti-symmetric with respect to the lines that divide vertically 
and horizontally the diagrams in half. In fact, this symmetry may be used to
construct all generators $\pceg$ as follows.\janote{sintetizar} Start out with $\tau_{0,\ldots,0}=1$
and begin to ``unfold'' the array of all $\pceg$'s coefficients $\taus$ by deciding
whether $\tau_{0,\ldots,0,1}$ be set to 1 or 0 (symmetric or anty-simmetric with 
respect to $\tau_{0,\ldots,0}$), then choose the values of $\{\tau_{0,\ldots,0,2}, 
\tau_{0,\ldots,0,3}\}$ to be symmetric or anti-symmetric with respect to 
$\{\tau_{0,\ldots,0},\tau_{0,\ldots,0,1}\}$; you need to keep this procedure 
for the remaining subsets with $2l$ coefficients $\taus$, with $l$ the cardinality of 
the previous subset, by setting the coefficients $\taus=0,1$ in order to be symmetric or
anti-symmetric to the previous subset. For example, let us illustrate the 
procedure to construct $\mcG_{(3,2)}$ represented in \Fref{fig:examplesPCE}(d).
One starts out choosing the value of $\tau_{(0,1)}$ to be anti-symmetric to $\tau_{(0,0)}$, then one chooses 
$\qty{\tau_{(0,2)},\tau_{(0,3)}}$ symmetric to $\qty{\tau_{(0,0)},\tau_{(0,1)}}$,
$\qty{\tau_{(1,0)},\tau_{(1,1)},\tau_{(1,2)},\tau_{(1,3)}}$ anti-symmetric 
to $\qty{\tau_{(0,0)},\tau_{(0,1)},\tau_{(0,2)},\tau_{(0,3)}}$, and finally
the last eight components $\tau_{\vec\alpha}$ anti-symmetric to the first 
eight.

\begin{align}
\tau_{(\vec\alpha)}
\end{align}

\begin{figure} % {{{
\centering
\includegraphics[width=\columnwidth]{oneQ_PCEGs-crop}
\caption{\janote{Propuesta de figura para mostrar que generadores 
y la a obedecen las mismas simetrías.}}
\end{figure} % }}}

\cpnote{Mucha lora y no se prueba nada, again}
The matrix $\san$ (and thus $a$) [see \eref{eq:san}] encodes all the information
of PCE generators $\pceg$ and, therefore, of all PCE quantum channels. 
The rows and columns of matrix $A$ also obey the symmetry previously discussed 
for PCE generators $\pceg$. It suffices to check that rows and columns of matrix $a$
in \Eref{eq:1} obey the aforementioned symmetry when they are arranged in a hypercubic 
$N$-dimensional array of order $4$.
As a consequence of $a$'s elements being $\pm1$, the power tensor $a^{\otimes N}$ will hold the symmetry of the 
rows and columns of $A$. \janote{falta prueba}
Therefore, this establishes a deep connection between the matrix $A$ and 
PCE quantum channels. It allows one to infer the PCE generators $\pceg$
by taking its corresponding coefficients $\{\taus\}$ as the rows or columns of $A$ 
after taking $-1\mapsto0$. In \appref{sec:2_qubits} we illustrate this program 
for 2-qubits PCE channels.
In short, it is sufficient to have the matrix $A$ to construct 
all PCE quantum channels.
%\janote{Aquí
%tendrá que ir una prueba que aún se está cocinando}

>>>>

We now discuss the existence of a generator set for all PCE quantum channels, 
how to label each of them uniquely as $\pceg$ according to its local action 
on every qubit in the system, and, finally, we will discuss a symmetry of
both PCE quantum channel generators and $A$, the matrix that diagonalizes
the Choi matrix $\mcD_N$ of a PCE map.

There exists a subset of PCE channels that generate 
all PCEs; the nature of these generators may be studied, as we
shall see, with the properties of the aforementioned vector space.
By standard theorems of linear algebra, any proper subspace $W$, see \Eref{eq:19}, 
can be extended to a maximal non-trivial subspace of dimension $2N-1$ by
adjoining appropriate additional basis elements. This can be done in
different ways. We therefore arrive to the set of maximal extensions of $W$,
where every maximal subspace corresponds to a PCE quantum channel that
preserves half of the Pauli components. The intersection of all the elements of
this set reduces to $W$ itself,
and since intersection of subspaces translates to composition of PCEs,
this implies that all PCEs can be obtained as
compositions of PCEs corresponding to maximally non-trivial subspaces, plus
the identity map. In other words, the set of PCEs that preserve half 
of the components plus the identity map, are a generator set for 
all PCE channels.
% and the complete vector space (corresponding to the identity channel),
% which we call generators. 
% \cpnote{Aca tenemos un semigrupo y no se si incluir a la identidad como 
% generador. De alguna manera es excepcional. Es problematico incluirla y es 
% problematico no incluirla.}
Consider \Fref{fig:examplesPCE}; subfigures (c), (d) and (e) represent nontrivial 
PCE generators and the composition of any two of them yield the PCE channel
corresponding to (b).

%\cpnote{Arreglar}
PCE generators may be characterized by its local action on every qubit 
in the system, and thus each of them be denoted as $\mcG_{\vec\alpha}$, 
with $\vec\alpha$ each of the $4^{n}$ different vector indices. For 1-qubit
the PCE generators are the identity map plus the channels that map the Bloch 
sphere to each Cartesian axis, as their associated vector subspaces are trivial
to prove and their dimension $2(1)-1=1$ is met. 
\janote{Here goes the bla bla for the discussion of the proof for the $N$-qubit
\pce{} generators}
Therefore, we can identify the index $\alpha_k$ in $\vec\alpha$ by the local action
of the PCE generator on the $k$-th qubit. 
See, for example, that the 2-qubits PCE generator represented in subfigure (c)
of \Fref{fig:examplesPCE} acts on the first qubit (first column) as mapping its
Bloch sphere to de $x$-axis, and on the second qubit (first row) acts an identity,
hence it is labeled $\mcG_{(1,0)}$.
See \Fref{fig:appex:twoQ_PCEGs} for the notation of all 2-qubit PCE generators.

%\cpnote{Mucha lora y no se prueba nada}
A reflection symmetry is identified for both PCE generators $\pceg$ and
the matrix $A$ that diagonalizes the Choi matrix $\mcD_N$ of a PCE map.
%PCE generators $\pceg$ obey a symmetry
%that defines uniquely every PCE generator. 
The symmetry for a PCE generator $\pceg$ is encoded in the 
$N$-dimensional array of all $\taus$, the Cartesian grid associated with it 
is either symmetric or anti-symmetric under reflection of the 
indices in the $k$-th position in $\vec\alpha$ as
\begin{align}
\alpha_{0}\mapsto\alpha_3,\quad\alpha_{3}\mapsto\alpha_0,
\quad\alpha_1\mapsto\alpha_2, \quad\alpha_2\mapsto\alpha_1.
\end{align}
\janote{Aquí iría la prueva de la simetría de reflexión que está en el horno aún.}
For instance, see that the 2-qubit PCE generators 
$\mcG_{(1,0)}$, $\mcG_{(3,2)}$ and $\mcG_{(2,2)}$
represented in subfigures (c), (d) and (e) of \Fref{fig:two:qubit:examples}, 
respectively, are either symmetric or anti-symmetric under reflection with 
respect to lines that divide the diagram in half vertically and horizontally. 
The same symmetry is identified for the rows and columns of $A$ if one 
labels each position in a row or column $A^{(i)}$ as $A^{(i)}_{\vec\alpha}$,
where $\vec\alpha$ is the same vector index used throughout in this manuscript.
\janote{Aquí habría que agregar la prueva de simetría para la $A$.}

\begin{figure} % {{{
\centering
\includegraphics[width=\columnwidth]{oneQ_PCEGs-crop}
\caption{
\janote{Propuesta de figura para mostrar que generadores 
y la a obedecen las mismas simetrías.}
}
\label{fig:A_and_PCEGs_connection}
\end{figure} % }}}

%\cpnote{Mucha lora y no se prueba nada, again}
Given that both matrix $\san$ and PCE generators $\pceg$ have the same 
symmetries, it is worth pointing out that $\san$ (and thus $a$) [see \eref{eq:san}] 
encodes all the information of PCE generators $\pceg$ and, therefore, of all PCE quantum channels. 
From the tensor power of matrix $a$ one can infer the $\{\taus\}$ of a PCE
generator $\pceg$ by taking any row or column $A^{(i)}$ and mapping 
all entries with $-1$ to $0$. In \Fref{fig:A_and_PCEGs_connection} we illustrate
this connection between $a$ and 1-qubit PCE generators.
In \appref{sec:2_qubits} we illustrate this program for 2-qubits PCE channels.
